---
title: "midterm"
author: "Yufei Zhao"
date: "4/23/2019"
output: html_document
---

```{r pkg, message = FALSE, warning = FALSE}
pacman::p_load(tidyverse, here, fs, glue)
```

# Part A: Data
```{r cache = TRUE} 
download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}


data <- map_df(c(15,16,17,18), ~ download_file(.x))

```

1. (5 points) Use the function above to download all the data for each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.
```{r}
data_1 <- data %>% 
  filter(`Student Group` %in%c("White", "Hispanic/Latino")) %>% 
  select(-contains("percent")) %>% 
  select(-contains("Proficient")) %>% 
  select(-c(13,14)) %>% 
  drop_na() %>% 
  gather(key = "level", value = "n", 9:12) %>% 
  separate(level, into=c("del1", "del2", "level")) %>% 
  select(-c("del1", "del2")) %>% 
  mutate(level = as.numeric(level))
head(data_1)
```

2 (2 points) Calculate the cumulative n for each school by student group, grade, and academic year. The result should look like the below. Hint, look at ?base::cumsum.
```{r}
data_2 <- data_1 %>% 
  group_by(`Academic Year`, `District ID`, `School ID`,Subject, `Student Group`, `Grade Level`) %>% nest() %>% 
  mutate(
    data = map(data, ~ arrange(., level)),
    data = map(data, ~ add_column(., cn = cumsum(.x$n)))
  ) %>% 
  unnest()
 head(data_2) 
```

3. (3 points) Reformat the data so it looks like the below, removing n and filling by cn. Remove rows that have missing data for either student group.
```{r}
data_3 <- data_2 %>% 
  select(-n) %>% 
  spread(key = `Student Group`, value = cn) %>% 
  drop_na()
head(data_3)
```

# Part B: Achievement gaps
```{r}
# install.packages("pracma")
gap <- function(data, ref, foc) {
    x <- data[[ref]]
    y <- data[[foc]]
    auc <- pracma::trapz(y / y[length(x)],
                         x / x[length(x)])
    sqrt(2)*qnorm(auc)
}

### Example
gap(data_3[1:4, ], "White", "Hispanic/Latino")
library(viridis)
```

1. (10 points) Estimate an achievement gap effect size for every school in the state that reported data on both student groups (i.e., using the data we created above), for each grade level in each academic year.
```{r}
data_4 <- data_3 %>% 
  group_by(`Academic Year`, `District ID`, District,`School ID`, School, `Grade Level`) %>% 
  nest() %>% 
  mutate(
    effect_size_white = map_dbl(data, ~ gap(.x, "White", "Hispanic/Latino")),
    effect_size_hispanic = map_dbl(data, ~ gap(.x, "Hispanic/Latino", "White"))
  )
```

2. (10 points) The plot below shows the achievement gap estimates for one school by grade in Ashland School District during the 2017-18 school year. Produce a similar plot to the below (noting the school, academic year, and school district) for each of the first 100 unique school/year/district combinations. Hint - you’ll want to use your effect size data from from the previous question, nest it, then apply slice(1:100). Note that the only reason I’m asking you to slice the data frame is just to reduce run time. In reality, you would do this for all school/year/district combinations.

```{r}
## test the plot appearance on a single plot

ggplot(data = data_5[[4]][[4]], aes(`Grade Level`,effect_size_white, fill = effect_size_white)) +
        geom_bar(stat = "identity", width = .5, alpha = 0.7) +
  coord_flip()  +
  theme_minimal() +
  labs(title = glue("Achievement Gap Estimates: ", unique(data_5[[4]][[4]]$School)),
       subtitle = "Students coded as White as compared to those coded as Hispanic/Latino",
       caption = glue("2017-18 School year, ", unique(data_5[[4]][[4]]$District), " School District, Oregon")) +
    scale_fill_viridis() +
  theme(legend.direction = "horizontal",
      legend.title=element_blank(),
      legend.justification=c(1,0),
      legend.position=c(1,-.45), plot.margin = margin(0.5, 0.5, 1.5, 0.5, "cm")
      )+
  xlab("Grade") +
  ylab("Effect Size")
  
```

```{r}
data_5 <- data_4 %>% 
  select(-data) %>% 
  filter(`Academic Year` == "2017-2018") %>% 
  separate(`Grade Level`, into = c("del", "Grade Level"), sep = " ", extra = "merge") %>% 
  select(-"del") %>%
  separate(District, into = c("District", "del"), sep = " SD", fill = "right") %>% 
  select(-"del") %>% 
  group_by(`Academic Year`, `District ID`,`School ID`) %>% 
  nest() %>% 
  slice(1:10) %>% 
  mutate(plots = map(data, ~ ggplot(data = .x, aes(`Grade Level`,effect_size_white, fill = effect_size_white)) +
        geom_bar(stat = "identity", width = .5, alpha = 0.7) +
  coord_flip()  +
    theme_minimal() +
  labs(title = glue("Achievement Gap Estimates: ", unique(data_5[[4]][[4]]$School)),
       subtitle = "Students coded as White as compared to those coded as Hispanic/Latino",
       caption = glue("2017-18 School year, ", unique(data_5[[4]][[4]]$District), " School District, Oregon")) +
    scale_fill_viridis() +
  theme(legend.direction = "horizontal",
      legend.title=element_blank(),
      legend.justification=c(1,0),
      legend.position=c(1,-.45), plot.margin = margin(0.5, 0.5, 1.5, 0.5, "cm")
      )+
  xlab("Grade") +
  ylab("Effect Size"))
  )
```

3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.
```{r}
project_dir <- path(here())
plots_dir <- path(project_dir, "plots")

data_6 <- data_5 %>% 
  mutate(plotnames =
           map_chr(data, ~ glue("2017-2018_", unique(.x$District), "_",unique(.x$School), ".png"))
         )

walk2(data_6$plots, data_6$plotnames, ~ ggsave(filename = .y, plot = .x, dpi = 300, path = plots_dir, width = 7, height = 8))
```

